A Bayesian ranking statistic to find \\high-mass black holes in LIGO data

### 2021-01-27 main.tex

@@

\begin{abstract}
The detection of intermediate mass black holes ( $10^2-10^6$~\msun) will shed light on the formation of supermassive black holes and thus galaxy formation. Although LIGO is sensitive to the merger of binary black holes with total masses up to $500$~\msun, only 1 of their 50 detections have a total mass $>100$~\msun. A possible explanation for the absence of intermediate mass events may be due to their misclassification as short-duration instrumental noise transients. Short-duration instrumental transients mimic the short-duration gravitational-wave signals from intermediate mass binary black hole mergers. Here we demonstrate that a search method utilising Bayesian inference could be a more sensitive tool for detecting high-mass binary black hole mergers (systems with a total mass $>55$~\msun) as compared to traditional match-filtering. We have applied this technique on the high-mass triggers during LIGO's second observing run to investigate the possibility of discovering new gravitational-wave signals from high mass black hole binaries, and to re-calculate the significance of high-mass candidate events. Although our search does not discover new candidate events, it does alter the significance of candidates identified by various search pipelines. 
\end{abstract}

@@

Since the 1970s, there has been an accumulation of evidence for stellar-mass and supermassive black holes. In April 2019, the Event Horizon Telescope provided the first visual evidence of the supermassive black hole M87~\cite{m87photo}. As of November 2020, the LIGO Scientific Collaboration has confirmed $\sim50$ binary black hole systems and listed numerous candidate events~\cite{GWTC1, GWTC2, gracedb}. Additionally, since the public release of LIGO's first and second observing run's data (O1 and O2), several external groups, such as \pycbc~\cite{pycbc_code} and a research team at the Institute for Advanced Study (IAS)~\cite{IAS0, IAS1, IAS2}, have independently searched and identified additional binary black hole gravitational wave candidates. These various discoveries have firmly established the existence of stellar-mass black holes, supermassive black holes and binary black hole systems.

Until recently, there was no definitive evidence for intermediate-mass black holes, the black holes that lie in between stellar-mass and supermassive black hole systems with masses between $10^2-10^6\ \msun$. This changed with the detection of GW190521, a unique gravitational wave event that lead to the formation of a black hole with a mass  $142\ \msun$, the first confirmed direct discovery of an intermediate mass black hole~\cite{gw190521}. Although this is the first gravitational wave that has lead to the discovery of a black hole with a mass in the intermediate range, ground based gravitational wave detectors are sensitive to gravitational waves from even more massive systems, up to a systems with a total mass of $400\ \msun$. \av{double check the following -- is this the most recent rate quote?} Additionally, gravitation waves from systems with masses greater than $100\ \msun$ systems should occur at a rate of $0-10\text{yr}^{-1}$~\cite{fregeau2006imbhbRatePrediction, mandel2008rates,rodriguez2015bbhRatePredictions,ligo_imbh_search}. Hence, we should have detected gravitational waves from more intermediate-mass systems, and potentially one event from O2.

However, even after conducting a targeted matched-filter based search for gravitational waves from intermediate-mass black holes in O2's data, the largest total mass detected so far is approximately $80\ \msun$~\cite{GWTC1,imbhbSearch2014, ligo_imbh_search}. A possible explanation for the absence of intermediate-mass events may be due to their misclassification as short-duration instrumental noise transients known as glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. These glitches can mimic astrophysical signals and hence decrease the significance of true gravitational wave events.

@@

Additionally, Bayesian odds can include comprehensive astrophysical information in its calculation. For example, a Bayesian odds to help detect gravitational-wave candidates can incorporate information on if an event's binary system was precessing and if the signal contains higher-order modes and is coherent amongst various detectors. Because of the possibility of Bayesian methods to incorporate detailed physical information about a gravitational wave signal, the LIGO Scientific Collaboration uses Bayesian methods to determine the source parameters of gravitational wave events~\cite{abbott2016ligo, GWTC1}. This paper demonstrates that the power of Bayesian methods used in parameter estimation can also discriminate between coherent gravitational-wave signals, incoherent glitches, and Gaussian noise in the form of a Bayesian odds, utilised as a ranking statistic.

In this paper we utilise a Bayesian method, called the Bayesian Coherence Ratio \bcr~\cite{BCR1}, to rank the candidate gravitational wave signals from high-mass compact binary coalescences (systems with total masses in the range of $55-500\ \msun$) in the detector data recorded during O2. Although the \bcr, utilising bootstrap techniques, does not provide the true Bayesian odds, it provides a more astrophysical measure of the significance of candidate events than a traditional match-filter significance.

We find that (a) high-mass events reported in the GWTC-1, including GW170729 (an event with disputed \pastro amongst various search pipelines) have high significance; (b) high-mass events detected from the IAS group have differing levels of significance; and that (c) our ranking statistic does not identify any intermediate mass black holes, but does identify an unreported stellar mass binary black hole candidate, GWC170222.

The remainder of this paper is structured as follows. We outline our methods, including details of the \bcr and the retrieval of our candidate events in Section~\ref{sec:method}. We present details on the implementation of our analysis in Section~\ref{sec:Analysis}. Finally we present our results in Section~\ref{sec:Results}, and discuss these results in the context of the significance of gravitational wave candidates in Section~\ref{sec:Conclusion}.

\section{Method\label{sec:method}}
The gravitational wave community uses Bayesian inference to perform parameter estimation and model selection. In this work, we utilise Bayesian inference to calculate the significance of high-mass candidate events in O2 by using the \bcr as a ranking statistic, taking a step forward to building a unified Bayesian framework to search for candidates and estimate their parameters.

@@

\footnotetext{\textit{Triggers} are points in time when the matched-filter SNR is greater than a threshold value for a given gravitational wave template~\cite{pycbc_og1}.}

@@

Whenever a local maximum of $\pycbcstat > \snr_\text{T}$, where $\snr_\text{T}$ is some threshold value, the search pipeline produces a single-detector trigger associated with the detector and time where the apparent signal in the data has its merger~\cite{pycbc_og6}.

For \pycbc to consider a trigger to be a \textit{candidate trigger}, a trigger from astrophysical origins, the trigger must be observed between detectors with the same template and a time of arrival difference less than the gravitational-wave travel time~\cite{pycbc_og1}. To test its search, \pycbc also conducts searches for \textit{simulated triggers}, artificial triggers manufactured by injecting signals into the detector data. These simulated signal studies provide \pycbc with metrics on its search's sensitivity. Finally, to quantify the statistical significance of candidate events, \pycbc artificially constructs a \textit{background trigger} set to compare against the candidate events. These background triggers are coherent signal-free events, constructed by applying relative offsets, or time-slides, between the data of different detectors~\cite{pycbc_og6}. Note that the time-slides to generate the background triggers are greater than the gravitational-wave travel time between detectors to ensure that the background triggers are not of astrophysical origins.

@@

After calculating the \bcr for a set of background triggers and simulated triggers from as stretch of detector-data (a data chunk), we can compute probability distributions for the background and simulated triggers, $p_b(\bcr)$ and $p_s(\bcr)$. We expect the background trigger and simulated signal \bcr values to favour the incoherent glitch and the coherent signal hypothesis, respectively. Ideally these distributions that represent two unique populations should be distinctly separate and have no overlap in their \bcr values. The prior odds parameters $\alpha$ and $\beta$ from Eq.~\ref{eq:bcr} help separate the two distributions. Altering $\alpha$ translates the \bcr probability distributions while adjusting $\beta$ spreads the distributions. Although Bayesian hyper-parameter estimation can determine the optimal values for $\alpha$ and $\beta$, an easier approach is to adjust the parameters for each data chunk's \bcr distribution. In this study, we tune $\alpha$ and $\beta$ to maximally separate the \bcr distributions for the background and simulated triggers.

@@

We expect the background trigger \bcr values to favour the incoherent glitch hypothesis (the null hypothesis). Candidate event \bcr values will either be statistically insignificant compared to the background triggers, implying the candidate is a glitch, or statistically significant to the background distribution, indicating the possible presence of an astrophysical signal. To quantify the level of significance we can calculate a false alarm probability with trial factors \fap for the candidate events. The \fap tells us how probable it is for the candidate event to be falsely identified as a signal.

@@

\subsection{Acquisition of triggers}
Advanced LIGO's second observing run O2 lasted $38$ weeks~\cite{GWOSC}. The software package, \pycbc~\cite{pycbc_code}, processed the O2 data in 22 time-frames (approximately 2 weeks for one time-frames) and found several gravitational wave events and numerous gravitational wave candidates~\cite{pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6}. Some candidate events were vetoed to be glitches, while others were rejected due to their low significance. The data is divided into these time-frames because the detector's sensitivity does not stay constant throughout 8 month long observing period.

@@

The most computationally intensive step during Bayesian inference is the evaluation of the likelihood $\mathcal{L}(d_i|\template)$. To accelerate our analysis, we use a likelihood that explicitly marginalises over coalescence time, phase at coalescence, and luminosity distance (Eq.~80 from~\citet{intro_to_gw_bayes}). While this marginalised likelihood reduces the run time without introducing errors to our evidence evaluation, it does not generates samples for the marginalised parameters. However, these parameter samples can be calculated as a post-processing step~\cite{intro_to_gw_bayes}.

We set the priors $\pi(\parameters|\mathcal{H}_S)$ and $\pi(\parameters|\mathcal{H}_G)$ to be identical. These priors restrict signals with mass parameters in the ranges presented in Table~\ref{tab:parameters}. The spins are aligned over a uniform range for the dimensionless spin magnitude from $\left[0,1\right]$. The luminosity distance prior assigns probability uniformly in comoving volume, with an upper cutoff of $5\ \text{Gpc}$. The full list of the priors, along with their shapes, limits and boundary conditions can be found in Table~\ref{tab:priors}.

@@

To generate the PSD, we take 31 neighbouring, off-source, non-overlapping,  4-second  segments of time-series data prior to the data segment $d_i$ being analysed. Off-soure data is used to avoid the inclusion of a signal in the PSD calculation. A Tukey window with a roll off of 0.2-seconds is applied to each data segment to suppress spectral leakage after which the segments are fast-Fourier transformed and median-averaged to create a PSD~\cite{ligo_psd}. This method, like other PSD estimation methods, adds statistical uncertainties to the PSD~\cite{psd_student_t, chatziioannou2019noise}. To marginalise over the statistical uncertainty we use the median-likelihood presented by~\citet{psd_student_t} as a post-processing step and shift our Bayesian Evidence estimations closer to their true astrophysical values.

Finally, we neglect detector calibration uncertainty and acquire data from from the Gravitational Wave Open Science Center~\cite{GWOSC}. The data we use is the publicly accessible O2 strain data from the Hanford and Livingston detectors. To ensure the data is usable we verify that the analysis and PSD data are obtained when the detectors are  in ``Science Mode''. The data requisition and quality checks are conducted using \gwpy~\cite{gwpy}.

The run-time to calculate a single Bayesian evidence after using \dynesty with $1,000$ live points and $100$ walkers is usually between $1-12\ \text{hours}$ (where the run time is proportional to the SNR of the data segment). The creation of the PSDs and execution of the parameter estimation jobs are managed via \bilbypipe~\cite{bilby_pipe}.

\subsection{Assigning \pastro to candidate events}

After the calculating the \bcr for the entire set of high-mass background and simulated triggers, we calculate probability distributions $p_b(\bcr)$ and $p_s(\bcr)$ for each 2-week time-frame of O2 data. These distributions are used to obtain the `tuned' prior-odd $\alpha$ and $\beta$ values that maximise $D_{KL}(p_b|p_s)$ for each time-frame of data.

Finally, using the tuned prior odds the \bcr for the candidate events can be calculated. Figure~\ref{fig:bcrCdf} shows the \bcr distributions for the background triggers, simulated triggers and candidate events. The bulk of the background and simulated trigger distributions are separate but have a slight overlap due to some of the simulated signal's being very faint. This suggests that the \bcr can successfully distinguish signals from noise or glitches. The vertical lines in Figure~\ref{fig:bcrCdf} displays the \bcr for gravitational wave candidate events. On comparing the candidate event \bcr values with the background distribution, we can estimate \pastro values for the candidate events.

@@

\input{Data/results_table}

We analyse the $60\ 996$ background, $5\ 146$ simulated, and $25$ foreground triggers reported by \pycbc's search of the data from the second observing run, restricting our analysis to the triggers that fall within our Bayesian prior space as described in Section~\ref{sec:method}. In addition to these triggers, we also analyse events and candidate events reported by GWTC-1 and the IAS group (note that some of these were identified as candidates by the \pycbc search). In Table~\ref{tab:results}, we summarise the \pastrobcr, along with the \pastro of other pipelines for comparison. Note that although the various pipeline \pastro are not mathematically equivalent, by comparing pipeline \pastro values for a given event we can compare how significant each pipeline deems various candidates.

@@

In addition to the above confirmed gravitational wave events from GWTC-1, we have also analysed several candidate events discussed in GWTC-1, most of which have low \pastrobcr. For example, consider the candidate event 170412, assigned a $\pastro$ of $0.06$ by \gstlal and has a $\pastrobcr$ of $0.01$. This candidate was reported to be excess power caused due noise appearing non-stationary between 60-200 Hz~\cite{GWTC1}. This candidate acts as an example of how $\pastrobcr$ may be utilised to not only to detect signals but also to eliminate terrestrial noise sources.

\subsection{IAS Events}
Our analysis of the \av{double check that these fall in our search space!!} high-mass IAS events and candidates in O2 has resulted in three events (GW170202, GW170403, GW170425) with disfavourable $\pastrobcr<0.5$. Additionally, five events (GW170121, GW170304, GW170402, GW170727, GW170817A) and one candidate event (GWC170302) reported by IAS have $\pastrobcr>0.5$.

Event candidates GW170727 and GWC170302 are of particular interest, with \pastrobcr of 0.94 and 0.73. GW170727 was emitted from a black hole binary system with a source frame total mass $\approx 70\ \msun$. In addition to the high \pastrobcr reported by our study, \citet{IAS1} and \citet{pycbc_ogc_2} have also reported high \pastro values of 0.98 and 0.99. On the other hand, GWC170302 is a sub marginal-candidate reported by \cite{IAS1} with a \pastro of 0.45 appears to have a higher significance our analysis, resulting in a \pastrobcr of $0.73$.

\av{GW170202's component black holes have has source masses (23, 16) below that of our prior space which may be why it has a low \pastrobcr}

@@

\subsection{New Candidate Events}
Although no clear detections were made with the \bcr, a marginal-candidate 170222 has been discovered with a \pastrobcr of $0.53$. This candidate has its similar masses when compared to those of GWTC-1.

>>>>

% symbols
\DeclareSymbolFont{extraup}{U}{zavm}{m}{n}
\DeclareMathSymbol{\varheart}{\mathalpha}{extraup}{86}
\DeclareMathSymbol{\vardiamond}{\mathalpha}{extraup}{87}

@@

\begin{abstract}
The detection of intermediate-mass black holes ( $10^2-10^6$~\msun) will shed light on the formation of supermassive black holes and thus galaxy formation. Although LIGO is sensitive to the merger of binary black holes with total masses up to $400$~\msun, only 4 of their 50 detections have a total mass $>100$~\msun. A possible explanation for the absence of intermediate-mass events may be their misclassification as short-duration instrumental noise transients. Short-duration instrumental transients mimic the short-duration gravitational-wave signals from intermediate-mass binary black hole mergers. Here we demonstrate that a ranking statistic utilising Bayesian inference could be a more sensitive tool for detecting high-mass binary black hole mergers (systems with a total mass $>55$~\msun) compared to a traditional match-filter ranking statistic. We applied this technique on the high-mass triggers during LIGO's second observing run to investigate the possibility of discovering new gravitational-wave signals from high mass black hole binaries and re-calculate the significance of high-mass candidate events. Although our analysis does not discover new gravitational wave events, it does provide a Bayesian measure of various candidates and events. 
\end{abstract}

@@

Since the 1970s, there has been an accumulation of evidence for stellar-mass and supermassive black holes. In April 2019, the Event Horizon Telescope provided the first visual evidence of the supermassive black hole M87~\cite{m87photo}. As of January 2021, the LIGO Scientific Collaboration has confirmed $\sim50$ binary black hole systems and listed numerous candidate events~\cite{GWTC1, GWTC2, gracedb}. Additionally, since the public release of LIGO's first and second observing run's data (O1 and O2), several external groups, such as \pycbc~\cite{pycbc_code} and a research team at the Institute for Advanced Study (IAS)~\cite{IAS0, IAS1, IAS2}, have independently searched and identified additional binary black hole gravitational wave candidates. These various discoveries have firmly authenticated the existence of stellar-mass black holes, supermassive black holes and binary black hole systems.

Until recently, there was no definitive evidence for intermediate-mass black holes, the black holes whose masses lie between that of stellar-mass and supermassive black hole systems ($10^2-10^6\ \msun$). This changed with the detection of GW190521, a unique gravitational wave event that led to the formation of a black hole with a mass  $142\ \msun$, the first confirmed direct discovery of an intermediate-mass black hole~\cite{gw190521}. Although this is the first gravitational wave event leading to discovering a black hole with a mass in the intermediate range, ground-based gravitational wave detectors are sensitive to gravitational waves from even more massive systems to systems with a total mass of $400\ \msun$. \av{double check the following -- is this the most recent rate quote?} Additionally, gravitational waves from systems with masses greater than $100\ \msun$ systems should occur at a rate of $0-10\text{yr}^{-1}$~\cite{fregeau2006imbhbRatePrediction, mandel2008rates,rodriguez2015bbhRatePredictions,ligo_imbh_search}. Hence, LIGO might have gravitational wave events from more intermediate-mass systems hidden in its data, potentially even in O2 data.

However, even after conducting a targeted matched-filter based search for gravitational waves from intermediate-mass black holes in O2's data, the heaviest total mass detected so far is approximately $80\ \msun$~\cite{GWTC1,imbhbSearch2014, ligo_imbh_search}. A possible explanation for the absence of intermediate-mass events may be due to their misclassification as short-duration instrumental noise transients known as glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. These glitches can mimic astrophysical signals and hence decrease the significance of true gravitational wave events.

@@

Additionally, Bayesian odds can include comprehensive astrophysical information in its calculation. For example, a Bayesian odds to help detect gravitational-wave candidates can incorporate information on if an event's binary system was precessing and if the signal contains higher-order modes and is coherent amongst various detectors. Because of the possibility of Bayesian methods to incorporate detailed physical information about a gravitational wave signal, the LIGO Scientific Collaboration uses Bayesian methods to determine the source parameters of gravitational-wave events~\cite{abbott2016ligo, GWTC1}. This paper demonstrates that the power of Bayesian methods used in parameter estimation can also discriminate between coherent gravitational-wave signals, incoherent glitches, and Gaussian noise in the form of a Bayesian odds, utilised as a ranking statistic.

In this paper, we utilise a Bayesian method, called the Bayesian Coherence Ratio \bcr~\cite{BCR1}, to rank the candidate gravitational wave signals from high-mass compact binary coalescences (systems with total masses in the range of $55-500\ \msun$) in the detector data recorded during O2. Although the \bcr, utilising bootstrap techniques, does not provide the true Bayesian odds, it provides a more astrophysical measure of candidate events' significance than a traditional match-filter significance.

We find that (a) high-mass events reported in the GWTC-1, including GW170729 (an event with disputed \pastro amongst various search pipelines) have high significance; (b) high-mass events detected from the IAS group have differing levels of significance; and that (c) our ranking statistic does not identify any intermediate-mass black holes, but does identify an unreported stellar-mass binary black hole candidate, 170222.

The remainder of this paper is structured as follows. We outline our methods, including details of the \bcr and the retrieval of our candidate events in Section~\ref{sec:method}. We present details on the implementation of our analysis in Section~\ref{sec:Analysis}. Finally, we present our results in Section~\ref{sec:Results}, and discuss these results in the context of the significance of gravitational wave candidates in Section~\ref{sec:Conclusion}.

\section{Method\label{sec:method}}
The gravitational wave community uses Bayesian inference to perform parameter estimation and model selection. In this work, we utilise Bayesian inference to calculate the significance of high-mass candidate events in O2 by using the \bcr as a ranking statistic, taking a step toward building a unified Bayesian framework to search for candidates and estimate their parameters.

@@

\footnotetext{\textit{Triggers} are points in time when the matched-filter SNR is larger than a threshold value for a given gravitational wave template~\cite{pycbc_og1}.}

@@

Whenever a local maximum of $\pycbcstat > \snr_\text{T}$, where $\snr_\text{T}$ is some threshold value, the \pycbc search pipeline produces a single-detector trigger associated with the detector and time where the apparent signal in the data has its merger~\cite{pycbc_og6}.

For \pycbc to consider a trigger to be a \textit{candidate trigger}, a trigger that may be from astrophysical origins, the trigger must be observed between detectors with the same template and a time of arrival difference less than the gravitational-wave travel time~\cite{pycbc_og1}. To test its search, \pycbc also conducts searches for \textit{simulated triggers}, artificial triggers manufactured by injecting signals into the detector data. These simulated signal studies provide \pycbc with metrics on its search's sensitivity. Finally, to quantify the statistical significance of candidate events, \pycbc artificially constructs a \textit{background trigger} set to compare against the candidate events. These background triggers are coherent signal-free events, constructed by applying relative offsets, or time-slides, between the data of different detectors~\cite{pycbc_og6}. The time-slides to generate the background triggers are greater than the gravitational-wave travel time between detectors to ensure that the background triggers are not of astrophysical origins.

@@

After calculating the \bcr for a set of background triggers and simulated triggers from as stretch of detector-data (a data chunk), we can compute probability distributions for the background and simulated triggers, $p_b(\bcr)$ and $p_s(\bcr)$. We expect the background trigger and simulated signal \bcr values to favour the incoherent glitch and the coherent signal hypothesis, respectively. Ideally, these distributions representing two unique populations should be distinctly separate and have no overlap in their \bcr values. The prior odds parameters $\alpha$ and $\beta$ from Eq.~\ref{eq:bcr} help separate the two distributions. Altering $\alpha$ translates the \bcr probability distributions while adjusting $\beta$ spreads the distributions. Although Bayesian hyper-parameter estimation can determine the optimal values for $\alpha$ and $\beta$, an easier approach is to adjust the parameters for each data chunk's \bcr distribution. In this study, we tune $\alpha$ and $\beta$ to maximally separate the \bcr distributions for the background and simulated triggers.

@@

We expect the background trigger \bcr values to favour the incoherent glitch hypothesis (the null hypothesis). Candidate event \bcr values will either be statistically insignificant compared to the background triggers, implying the candidate is a glitch, or statistically significant to the background distribution, indicating the possible presence of an astrophysical signal. To quantify the significance level, we can calculate a false alarm probability with trial factors \fap for the candidate events. The \fap is the probability that a candidate event originating from a non-astrophysical source can be is falsely identified as a signal.

@@

\subsection{Acquisition of triggers}
Advanced LIGO's second observing run O2 lasted $38$ weeks~\cite{GWOSC}. The software package, \pycbc~\cite{pycbc_code}, processed the O2 data in 22 time-frames (approximately 2 weeks for one time-frames) and found several gravitational wave events and numerous gravitational wave candidates~\cite{pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6}. Some candidate events were vetoed to be glitches, while others were rejected due to their low significance. The data is divided into these time-frames because the detector's sensitivity does not stay constant throughout the eight-month-long observing period.

@@

The most computationally intensive step during Bayesian inference is evaluating the likelihood $\mathcal{L}(d_i|\template)$. To accelerate our analysis, we use a likelihood that explicitly marginalises over coalescence time, phase at coalescence, and luminosity distance (Eq.~80 from~\citet{intro_to_gw_bayes}). While this marginalised likelihood reduces the run time without introducing errors to our evidence evaluation, it does not generates samples for the marginalised parameters. However, these parameter samples can be calculated as a post-processing step~\cite{intro_to_gw_bayes}.

We set the priors $\pi(\parameters|\mathcal{H}_S)$ and $\pi(\parameters|\mathcal{H}_G)$ to be identical. These priors restrict signals with mass parameters in the ranges presented in Table~\ref{tab:parameters}. The spins are aligned over a uniform range for the dimensionless spin magnitude from $\left[0,1\right]$. The luminosity distance prior assigns probability uniformly in comoving volume, with an upper cutoff of $5\ \text{Gpc}$. The full list of the priors, along with their shapes, limits and boundary conditions are documented in Table~\ref{tab:priors}.

@@

We take 31 neighbouring, off-source, non-overlapping,  4-second  segments of time-series data before the analysis data segment $d_i$ to generate the PSD. We use off-source to avoid the inclusion of a signal in the PSD calculation. A Tukey window with 0.2-second roll-offs is applied to each data segment to suppress spectral leakage after which the segments are fast-Fourier transformed and median-averaged to create a PSD~\cite{ligo_psd}. Like other PSD estimation methods, this method adds statistical uncertainties to the PSD~\cite{psd_student_t, chatziioannou2019noise}. To marginalise over the statistical uncertainty, we use the median-likelihood presented by~\citet{psd_student_t} as a post-processing step and shift our Bayesian Evidence estimations closer to their astrophysical values.

Finally, we neglect detector calibration uncertainty and acquire data from from the Gravitational Wave Open Science Center~\cite{GWOSC}. The data we use is the publicly accessible O2 strain data from the Hanford and Livingston detectors, recorded while the detectors are in ``Science Mode''. We obtain the data using \gwpy~\cite{gwpy}.

The run-time to calculate a single Bayesian evidence after using \dynesty with $1,000$ live points and $100$ walkers is usually between $1-12\ \text{hours}$ (where the run time is proportional to the SNR of the data segment). \bilbypipe orchestrates the creation of the PSDs and execution of the parameter estimation jobs~\cite{bilby_pipe}.

\subsection{Assigning \pastro to candidate events}
After the calculating the \bcr for the entire set of high-mass background and simulated triggers, we calculate probability distributions $p_b(\bcr)$ and $p_s(\bcr)$ for each 2-week time-frame of O2 data. These distributions are used to obtain the `tuned' prior-odd $\alpha$ and $\beta$ values that maximise $D_{KL}(p_b|p_s)$ for each time-frame of data.

Finally, using the tuned prior odds the \bcr for the candidate events can be calculated. Figure~\ref{fig:bcrCdf} shows the \bcr distributions for the background triggers, simulated triggers and candidate events. The bulk of the background and simulated trigger distributions are separate but slightly overlap due to some of the simulated signal's being very faint. The separation suggests that the \bcr can successfully distinguish signals from noise or glitches. The vertical lines in Figure~\ref{fig:bcrCdf} displays the \bcr for gravitational wave candidate events. On comparing the candidate event \bcr values with the background distribution, we can estimate \pastro values for the candidate events.

@@

\input{Data/results_table}
We analyse the $60\ 996$ background, $5\ 146$ simulated, and $25$ foreground triggers reported by \pycbc's search of the data from the second observing run, restricting our analysis to the triggers that fall within our Bayesian prior space as described in Section~\ref{sec:method}. In addition to these triggers, we also analyse events and candidate events reported by GWTC-1 and the IAS group (note that some of these were identified as candidates by the \pycbc search). In Table~\ref{tab:results}, we summarise the \pastrobcr, along with the \pastro of other pipelines for comparison. Note that although the various pipeline \pastro are not mathematically equivalent, by comparing pipeline \pastro values for a given event, we can compare how significant each pipeline deems various candidates.

@@

In addition to the above confirmed gravitational wave events from GWTC-1, we have also analysed several candidate events discussed in GWTC-1, most of which have low \pastrobcr. For example, consider the candidate event 170412, assigned a $\pastro$ of $0.06$ by \gstlal and has a $\pastrobcr$ of $0.01$. This candidate was reported to be excess power caused due noise appearing non-stationary between 60-200 Hz~\cite{GWTC1}. This candidate acts as an example of how $\pastrobcr$ may be utilised  to eliminate terrestrial noise sources, in addition to detecting signals

\subsection{IAS Events}
Our analysis of the \av{double check that these fall in our search space!!} high-mass IAS events and candidates in O2 has resulted in three events with disfavourable $\pastrobcr<0.5$ (GW170202, GW170403, GW170425). On the other hand, five events (GW170121, GW170304, GW170402, GW170727, GW170817A) and one candidate event (170302) reported by IAS have $\pastrobcr>0.5$.

GW170727 and 170302 are of particular interest, with \pastrobcr of 0.94 and 0.73. GW170727 was emitted from a black hole binary system with a source frame total mass $\approx 70\ \msun$. In addition to the high \pastrobcr reported by our study, \citet{IAS1} and \citet{pycbc_ogc_2} have also reported high \pastro values of 0.98 and 0.99, making it a viable gravitational-wave event candidate. Similarly, the sub-marginal-candidate 170302 reported by \cite{IAS1} with a \pastro of 0.45 appears to have a higher significance from our analysis, resulting in a \pastrobcr of $0.73$.

\av{GW170202's component black holes have source masses (23, 16) below that of our prior space which may be why it has a low \pastrobcr}

@@

\subsection{New Candidate Events}
Although no clear detections are made with the \bcr, a marginal-candidate 170222 has been discovered with a \pastrobcr of $0.53$. This candidate has its similar masses when compared to those of GWTC-1. The other candidates with \pastrobcr
